Moral Foundations of U.S. Political News Organizations

### 2019-01-21 pilot_manuscript.tex
updated thesis

\documentclass[,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Moral Foundations of U.S. Political News Organizations},
            pdfauthor={William E. Padfield~\& Erin M. Buchanan, Ph.D.},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

Something clever to start

@@

Jonathan Haidt and Jesse Graham formulated Moral Foundations Theory as a
method by which to capture the entirety of humans' moral domain. The
researchers argued older theories of moral psychology were focused
primarly on issues of justice, fairness, and caring - indvidually
focused foundations of morality that align with the beliefs of political
liberals ({\textbf{???}}). In other words, moral psychology ignored the
valid moral foundations of conservatives. Moral Foundations Theory (MFT)
holds that people's moral domain can be mapped by quantifying their
endorsement of five moral foundations: \emph{harm/care},
\emph{fairness/reciprocity}, \emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity} ({\textbf{???}}).

The researchers settled on these specific foundations after the
completion of a literature survey of research in anthropology and
evolutionary psychology ({\textbf{???}}).The first two foundations
(\emph{harm/care} and \emph{fairness/reciprocity}) are termed the
\enquote{individualizing foundations,} as they are centered on the
concerns of individuals rather than groups. \emph{Harm/care} represents
an endorsement of compassion and kindness, while opposing cruelty and
harm. \emph{Fairness/reciprocity} represents concerns centered on
guaranteeing individual rights as well as justice and equality among all
people. The other three foundations (\emph{ingroup/loyalty},
\emph{authority/respct}, and \emph{purity/sanctity}) are the
\enquote{binding} foundations, owing to their focus on group-related
concerns, rather than those of individuals. \emph{Ingroup/loyalty}
represents endorsements of patriotism and heroism and discourages
nonconformity and dissent. \emph{Authority/respect} represents an
endorsement of social hierarchies and traditions while denigrating
disobedience. Finally, \emph{purity/sanctity} represents concerns
regarding chastity and piety, while discouraging vices and indulgences,
including lust, avarice, and gluttony ({\textbf{???}}). Liberals tend to
endorse the individualizing foundations more than conservatives.
Endorsements for the three binding foundations, however, are lower than
conservatives. Conservatives, on the other hand, tend to endorse all
five foundations equally, implying they base judments (at least
partially) on each foundation ({\textbf{???}}).

@@

In order to capture language's role in moral and political reasoning,
({\textbf{???}}) formulated the Moral Foundations Dictionary (MFD) in
order to capture moral reasoning and justification as used in speech and
text. The MFD is composed of 259 words, with around 50 words assigned to
each of the five foundations. The researchers created a preliminary list
of words that they believed would be associated with the five
foundations. Then, using the Linguistic Inquiry and Word Count (LIWC;
{\textbf{???}}) computer program, they analyzed transcripts of liberal
and conservative Christian sermons in order to obtain frequencies of the
occurrence of words from the researchers' initial list. The researchers
manually checked the results from LIWC in order to make sure the results
make sense given the contexts and rhetorical devices used in the
sermons. Similar to previous research on Moral Foundations Theory,
liberal ministers used \emph{harm}, \emph{fairness}, and \emph{ingroup}
words more often than conservative ministers. Conversely, conservative
ministers used \emph{authority} and \emph{purity} words more often than
liberal ministers. However, conservative ministers did not use
\emph{ingroup/loyalty} words more than liberals. Rather, liberal
ministers used words pertaining to \emph{ingroup/loyalty}, but in
contexts that promote rebellion and independence - causes
\emph{opposite} to positive endorsements of that foundation
({\textbf{???}}).

To this point, most text analysis utilizing the Moral Foundations
Dictionary operationalizes endorsement of any one of the foundations as
percent occurence of words from the foundation's respective word list.
As such, most analyses assume that zero percent occurence is indicative
of no endorsement, while any non-zero percent occurence indicates
endorsement of the foundation. This may not be sufficient in describing
the true nature of the writer or speaker's endorsement of one of the
sets of moral intuitions. A quick glance at the MFD words for
\emph{harm/care} reveals the presence of words that are more closely
associated with universally accepted conceptions of \emph{harm} over
\emph{care} and vice-versa ({\textbf{???}}). For example, the word
\enquote{cruel} has relatively negative connotations compared to
\enquote{benefit.} For the \emph{harm/care} foundation, it is
conceivable that use of the word \enquote{cruel} might indicate a
greater attentional focus of the idea of \emph{harm} rather than
\emph{care}.

For \emph{harm/care}, the definition of the foundation as well as its
name clearly distinguishes between two somewhat opposite sides of an
attentional continuum, with \emph{harm} on the negative end and
\emph{care} on the positive side. In other words, the entries in the MFD
for \emph{harm/care} have somewhat clear positive and negative valences.
The same pattern can be seen in the MFD entries for the other four
foundations. \emph{Purity/sanctity} features words that likely have a
negative valence to most observers, including \enquote{disease}" and
\enquote{trash,} along with more positive words, including
\enquote{right} and \enquote{sacred} ({\textbf{???}}). This, however,
brings up other questions regarding the definition and names of the
other four foundations apart from \emph{harm/care}:
\emph{fairness/reciprocity}, \emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity}. The latter four
foudnations have names that are harder to understand as a valence
continuum, as the concepts in the names are more similar, even to the
point of being virtually synonymous in the case of
\emph{fairness/reciprocity}.

When considering the issue of positive versus negative valence in MFD
words, the question of how texts are analyzed vis-a-vis the MFD remains.
How can raw percentage of MFD word occurence capture the valence and
focus of the writer or speaker? If 2\% of a politician's speech features
positive words (i.e. \enquote{benefit} and \enquote{defend}) from the
MFD \emph{harm/care} list, how can researchers be sure the level and
nature of the speaker's \enquote{endorsement} of the foundation equals
that of another politician whose speech contained negatively connoted
MFD words from the \emph{harm/care} list? They would have equal
endorsements as far as the numbers are concerned, but the words used and
focus given are on opposite sides of the \emph{harm/care} spectrum.

This issue is compounded by the fact the Moral Foundations Questionnaire
(MFQ) and its subscales assume endorsement lies on a continuum. One item
under the \emph{fairness/reciprocity} judgment subscale reads,
\enquote{Justice is the most important requirement for a society}
({\textbf{???}}). The survey respondent must select a number on a scale
from 1 to 6 indicating responses spanning \enquote{strongly disagree} at
1 to \enquote{strongly agree} at 6. While the scales in the MFQ do not
represent true valence as it pertains to individual words, it does allow
for a greater degree of specificity in terms of an individual's
endorsement of a particular moral foundation. When a respondent selects
a 4 for the aforementioned MFQ statement, they clearly are indicating
they \enquote{slightly agree} with the statement ({\textbf{???}}). This
specificity is not present in most analyses involving the MFD and
percent occurence, unless they also take into account the valence of the
words used in the text or speech of interest.

@@

Borrowing from Osgood's work in the 1950s, ({\textbf{???}}) recognized
valence as one of three related dimensions comprising emotion when
developing their Affective Norms for English Words (ANEW). As mentioned
before, \enquote{valence,} the first dimension, denotes the pleasantness
of a given word. \enquote{Arousal,} the second dimension, describes the
stimulating nature of a word. Lastly, \enquote{dominance} or
\enquote{control} describes the extent to which a word makes one feel in
or out of control ({\textbf{???}}). The researchers developed ANEW by
presenting participants with a list of 100-150 words and asking for them
to rate the word on all three dimensions using the Self-Assessment
Mannikin (SAM), which allows ratings along either a nine-point scale
when using traditional paper instruments or a twenty-point scale when
using a computerized version.

Participants saw the stimulus word and responded on each scale. The
valence scale featured a smiling figure at one end (representing
pleasantness) and a frowning figure at the other end (for
unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at
one end with a sleepy figure at the other, representing stimulating and
unstimulating respectively. Finally, the control scale featured a large
figure, indicating the highest degree of control, at one end and a small
figure, indicating a lack of control, at the other end ({\textbf{???}}).
The end result of this procedure yielded affective norms along the three
dimensions for 1,040 English words ({\textbf{???}}). ANEW represented an
important first step in establishing affective norms for large numbers
of English words. However, later researchers found the 1,040-word list
to be limiting for a language consisting of thousands of words.

({\textbf{???}}) exponentially lengthened the list of words with
affective norms to 13,915 English lemmas, the base forms of words
without inflection (i.e. \enquote{watch} rather than \enquote{watched}
and \enquote{watching}). The researchers recognized the importance of
affective norms in several areas of study, including emotion, language
processing, and memory ({\textbf{???}}). They argue the list of words
included in ANEW is sufficient for small-scale factorial research
designs, but the list is \enquote{prohibitively small} for larger-scale
\enquote{megastudies} that are common in psycholinguistic research today
({\textbf{???}}).

In order to source a large number of lemmas for affective ratings, the
researchers drew from several validated sources. These include the
30,000 lemmas with age-of-acquisition (average age at which a particular
word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and
Brysbaert as well as the content lemmas from the SUBTLEX-US corpus
consisiting of subtitles from various forms of visual media
({\textbf{???}}). This resulted in the final list of 13,915 lemmas.
Lists of 346-350 words were presented to participants recruited through
the Amazon Mechanical Turk subject pool. Particpants rated the words
along one of the three dimensions, unlike the ANEW project in which
participants rated each word along all three dimensions at once. The
researchers used a nine-point scale similar to the one used by
({\textbf{???}}) when collecting ratings for ANEW ({\textbf{???}}).

The researchers noted several points of interest upon observing ratings.
First, they found that valence and dominance ratings had a negative
skew, indicating more words elicited feelings of happiness and control
than their respective opposites. Also, when examining the relationship
between valence and arousal ratings, the researchers found a U-shaped
relationship. This indicates words with high degrees of positivity and
negativity elicited higher arousal ({\textbf{???}}). These observations
along with the now-greatly expanded list of affective norms has been
applied to several lines of inquiry in psycholinguistics.

({\textbf{???}}) utilized the new affective norms list in order to
investigate the validity of the Pollyanna hypothesis, or the prevalence
of a generally optimistic outlook in humans as reflected in langauge.
The researchers were able to conclude the existence of a greater number
of positive-valence English words in the list of 13,915 lemmas.
Additionally, after observing token frequency in a number of text
corpora, including SUBTLEX-US, COCA, BNC, TASA, and HAL, the researchers
found that words with postive valence were also used more frequently
({\textbf{???}}). While the researchers concede the possiblility of an
acquiensence bias in ratings as a possible explanation for the observed
positivity bias, this investigation represents one application of the
({\textbf{???}}) list in emotional studies.

One cognition-based study investigates the relationship between emotion
and response latencies in word recognition. ({\textbf{???}}) sought to
use the ({\textbf{???}}) norms to fill in the knowledge gaps regarding
variance in word recognition. The researchers drew several conclusions
regarding emotion and word recognition (specifically in naming and
lexical decision tasks). First, ({\textbf{???}}) found higher response
latencies in negative-valence words, lower response latencies in neutral
words, and even lower response latencies in words with positive valence.
The researchers also concluded that words causing higher arousal tend to
have higher response latencies than less-arousing words. They found
valence had a stronger effect on recognition than arousal (both effects
were independent, not interactive). They found an interaction between
emotion and word frequency such that valence and arousal are more
effective on lower frequency words than high frequency words. Finally,
({\textbf{???}}) found a greater effect for lexical decision tasks than
for naming tasks. This research serves as further evidence that the
({\textbf{???}}) list can be used for research inquiries both within and
without the field of psycholinguistics.

In the present studies, the researchers used the 2013 Warriner list in
order to denote the valence of the words appearing in the news articles
scraped from the internet. Valence was considered as another independent
variable and its relationship with the words comprising the Moral
Foundations Dictionary were of chief interest to the researchers. The
valence was used as a means to determine whether individiual words in
the MFD represented more positive aspects of their respective foundation
or if they denoted a more negative aspect of the foundation.
Incorporating word valence into a study involving the MFD is meant to
alleviate some of the issues regarding the aforementioned ambiguity
regarding the words in the Moral Foundations Dictionary.

@@

How you would run the data analysis

>>>>

\documentclass[,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Moral Foundations of U.S. Political News Organizations},
            pdfauthor={William E. Padfield~\& Erin M. Buchanan, Ph.D.},
            pdfkeywords={politics, morality, psycholinguistics},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\begin{document}
\maketitle

In the United States, today's media landscape affords consumers a multitude of options for obtaining political news. Since the advent of cable news networks and the World Wide Web in the last decades of the twentieth century, consumers have gained access to an ever-expanding menagerie of news sources, many of which can be called up via a simple click, touch, or swipe. Concurrent with this growth in available news sources, concerns regarding political bias in news reporting have entered public consciousness. For example, commentators argue that networks including Fox News Channel and MSNBC communicate political news from a conservative and liberal slant respectively. These purported biases have been a cause for concern given the potential for incomplete or inaccurate news reporting potentially resulting from these biases. Given the inherently moral nature of many political arguments and positions, bias in news reporting might manifest as differing moral appeals. Specifically, the use of differing moral language in political articles might be an indicator of political bias in news media.

Morality and ethics have been of interest to thinkers, academics, and philosophers since antiquity. Starting chiefly in the twentieth century, a scientific approach to humans' understanding of morality emerged under the domain of psychology. Theories attempting to explain the development and application of people's moral intuitions built the foundation for the subfield of moral psychology. As the field developed, however, considerable debate has taken place regarding operational definitions of \enquote{morality.} This remains an issue in the field in the twenty-first century as researchers attempt to infer moral and political leanings from text and speech.

@@

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as they mature. Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage (Haidt \& Graham, 2007). Similarly, Kohlberg found people typically start with a \enquote{pre-conventional} understanding of morality during infancy and progress through several steps until they reach a \enquote{post-conventional} ethics (Haidt \& Graham, 2007). People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice. From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality (Haidt \& Graham, 2007). This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women (Haidt \& Graham, 2007). She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. Gilligan thus asserted morality was built upon an alternative moral foundation: caring (Haidt \& Graham, 2007). This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring (Haidt \& Graham, 2007). While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

In their brief overview of the history of moral psychology, Graham, Haidt, and Nosek (2009) explained Shweder, Much, Mahapatra, and Park's objections to moral psychology as it stood in the late 1980s. Their criticism centered on the fact moral psychology concerned itself with issues regarding justice and individuals' rights. Such a system, they argued, did not account for moral concerns outside of the western world (Graham et al., 2009). Individually focused concerns can be grouped under an overarching \enquote{ethic of autonomy,} which was thought to be one of three ethics upon which humans base moral decisions. The other two ethics were the \enquote{ethic of community} (comprising one's duty to their family, tribe, etc.) and the \enquote{ethic of divinity}" - representing one's duty not to defile their God-given body and soul (Graham et al., 2009). In the 2000s, Haidt and Graham (2007) took this line of reasoning further in their assertion that moral psychology favored certain political ideologies over others.

Jonathan Haidt and Jesse Graham formulated Moral Foundations Theory as a method by which to capture the entirety of humans' moral domain (Haidt \& Graham, 2007). The researchers argued older theories of moral psychology were focused primarily on issues of justice, fairness, and caring - individually focused foundations of morality that align with the beliefs of political liberals (Haidt \& Graham, 2007). In other words, moral psychology ignored the valid moral foundations of conservatives. Moral Foundations Theory (MFT) holds that people's moral domain can be mapped by quantifying their endorsement of five moral foundations: \emph{harm/care}, \emph{fairness/reciprocity}, \emph{ingroup/loyalty}, \emph{authority/respect}, and \emph{purity/sanctity} (Haidt \& Graham, 2007).

The researchers settled on these specific foundations after the completion of a literature survey of research in anthropology and evolutionary psychology (Graham et al., 2011). The researchers attempted to locate virtues and morals corresponding to \enquote{evolutionary thinking.} For instance, the researchers cited Mauss' work on reciprocal gift-giving, which informed the establishment of the \emph{fairness/reciprocity} foundation. Additionally, evolutionary literature on disgust and its correlation to human behavior regarding food and sex informed the \emph{purity/sanctity} foundation (Graham et al., 2011). The researchers identified the five \enquote{top candidates} for the foundations of human cultures' morality (Graham et al., 2011).

The first two foundations (\emph{harm/care} and \emph{fairness/reciprocity}) are termed the \enquote{individualizing foundations,} as they are centered on the concerns of individuals rather than groups. \emph{Harm/care} represents an endorsement of compassion and kindness, while opposing cruelty and harm. \emph{Fairness/reciprocity} represents concerns centered on guaranteeing individual rights as well as justice and equality among all people. The other three foundations (\emph{ingroup/loyalty}, \emph{authority/respect}, and \emph{purity/sanctity}) are the \enquote{binding} foundations, owing to their focus on group-related concerns, rather than those of individuals. \emph{Ingroup/loyalty} represents endorsements of patriotism and heroism and discourages nonconformity and dissent. \emph{Authority/respect} represents an endorsement of social hierarchies and traditions while denigrating disobedience. Finally, \emph{purity/sanctity} represents concerns regarding chastity and piety, while discouraging vices and indulgences, including lust, avarice, and gluttony (Haidt \& Graham, 2007). Liberals tend to endorse the individualizing foundations more than conservatives. Conservatives, on the other hand, tend to endorse the binding foundations more than liberals. It should be noted, however, conservatives also tend to endorse all five foundations equally, implying they base moral judgments on all foundations (Graham et al., 2009).

@@

In order to capture language's role in moral and political reasoning, Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD) in order to capture moral reasoning and justification as used in speech and text. The MFD is composed of 259 words, with around 50 words assigned to each of the five foundations. The researchers created a preliminary list of words that they believed would be associated with the five foundations. Then, using the Linguistic Inquiry and Word Count (LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they analyzed transcripts of liberal and conservative Christian sermons in order to obtain frequencies of the occurrence of words from the researchers' initial list. The researchers manually checked the results from LIWC in order to make sure the results make sense given the contexts and rhetorical devices used in the sermons. The researchers offered the following example from a Unitarian sermon as a demonstration of ambiguous statements requiring human verification: \enquote{Don't let some self-interested ecclesiastical or government authority tell you what to believe, but read the Bible with your own eyes and open your heart directly to Jesus} (Graham et al., 2009). This sentence added to the \emph{authority/respect} total in LIWC's analysis, but it appears to suggest that one should reject authority in this context. The researchers eliminated this sentence from the \emph{authority/respect} raw count on account of this discrepancy between the use of authority-related words and the speaker's clear intentions (Graham et al., 2009).

Similar to previous research on Moral Foundations Theory, liberal ministers used \emph{harm}, \emph{fairness}, and \emph{ingroup} words more often than conservative ministers. Conversely, conservative ministers used \emph{authority} and \emph{purity} words more often than liberal ministers. However, conservative ministers did not use \emph{ingroup/loyalty} words more than liberals. Rather, liberal ministers used words pertaining to \emph{ingroup/loyalty}, but in contexts that promote rebellion and independence - causes \emph{opposite} to positive endorsements of that foundation (Graham et al., 2009).

To this point, most text analysis utilizing the Moral Foundations Dictionary operationalizes endorsement of any one of the foundations as percent occurrence of words in a given text from the foundation's respective word list. As such, most analyses assume that zero percent occurrence is indicative of no endorsement, while any non-zero percent occurrence indicates endorsement of the foundation. This operational definition may not be sufficient in describing the true nature of the writer or speaker's endorsement of one of the sets of moral intuitions. A quick glance at the MFD words for \emph{harm/care} reveals the presence of words that are more closely associated with universally accepted conceptions of \emph{harm} over \emph{care} and vice-versa (Graham et al., 2009). For example, the word \enquote{cruel} has relatively negative connotations compared to \enquote{benefit.} For the \emph{harm/care} foundation, it is conceivable that use of the word \enquote{cruel} might indicate a greater attentional focus of the idea of \emph{harm} rather than \emph{care}.

For \emph{harm/care}, the definition of the foundation, as well as its name, clearly distinguishes between two somewhat opposite sides of an attentional continuum, with \emph{harm} on the negative end and \emph{care} on the positive side. In other words, the entries in the MFD for \emph{harm/care} have somewhat clear positive and negative valences. The same pattern can be seen in the MFD entries for the other four foundations. \emph{Purity/sanctity} features words that likely have a negative valence to most observers, including \enquote{disease}" and \enquote{trash,} along with more positive words, including \enquote{right} and \enquote{sacred} (Graham et al., 2009). These dichotomies, however, bring up other questions regarding the definition and names of the other four foundations apart from \emph{harm/care}: \emph{fairness/reciprocity}, \emph{ingroup/loyalty}, \emph{authority/respect}, and \emph{purity/sanctity}. The latter four foundations have names that are harder to understand as a valence continuum, as the concepts in the names are more similar, even to the point of being virtually synonymous in the case of \emph{fairness/reciprocity}.

When considering the issue of positive versus negative valence in MFD words, the question of how texts are analyzed vis-a-vis the MFD remains. How can raw percentage of MFD word occurrence capture the valence and focus of the writer or speaker? If 2\% of a politician's speech features positive words (i.e. \enquote{benefit} and \enquote{defend}) from the MFD \emph{harm/care} list, how can researchers be sure the level and nature of the speaker's \enquote{endorsement} of the foundation equals that of another politician whose speech contained negatively connoted MFD words from the \emph{harm/care} list? They would have equal endorsements as far as the numbers are concerned, but the words used and focus given are on opposite sides of the \emph{harm/care} spectrum.

This issue is compounded by the fact the Moral Foundations Questionnaire (MFQ) and its subscales assume endorsement lies on a continuum. The Moral Foundations Questionnaire (MFQ), which was developed subsequent to the MFD, measures individuals' endorsements of each of the foundations using a six-point scale (Graham et al., 2011). The questionnaire is made up of judgment items and relevance items. Judgment items are phrased such that the respondent signals their agreement or disagreement with straightforward statements. An example of such a statement reads: \enquote{It can never be right to kill another human being} (Graham et al., 2011). Relevance items gauge the respondent's opinion regarding the importance of foundation-related concerns. For example, the respondent is directed to rate how important the following situation is to their sense of morals: \enquote{whether or not someone did something disgusting.} This example measures the relevance of the \emph{purity/sanctity} foundations. Each foundation has a judgment and relevance subscale, totaling 10 subscales for the MFQ (Graham et al., 2011).

The aforementioned ambiguity of the Moral Foundations Dictionary as an instrument becomes clearer upon closer examination of the items in the Moral Foundations Questionnaire. One item under the \emph{fairness/reciprocity} judgment subscale reads, \enquote{Justice is the most important requirement for a society} (Graham et al., 2011). The survey respondent must select a number on a scale from 1 to 6 indicating responses spanning \enquote{strongly disagree} at 1 to \enquote{strongly agree} at 6. While the scales in the MFQ do not represent true valence as it pertains to individual words, it does allow for a greater degree of specificity in terms of an individual's endorsement of a particular moral foundation. When a respondent selects a 4 for the aforementioned MFQ statement, they clearly are indicating they \enquote{slightly agree} with the statement (Graham et al., 2011). This specificity is not present in most analyses involving the MFD and percent occurrence, unless they also take into account the valence of the words used in the text or speech of interest.

@@

Borrowing from Osgood's work in the 1950s, Bradley and Lang (1999) recognized valence as one of three related dimensions comprising emotion when developing their Affective Norms for English Words (ANEW). As mentioned before, \enquote{valence,} the first dimension, denotes the pleasantness of a given word. \enquote{Arousal,} the second dimension, describes the stimulating nature of a word. Lastly, \enquote{dominance} or \enquote{control} describes the extent to which a word makes one feel in or out of control (Bradley \& Lang, 1999). The researchers developed ANEW by presenting participants with a list of 100-150 words and asking for them to rate the word on all three dimensions using the Self-Assessment Mannikin (SAM), which allows ratings along either a nine-point scale when using traditional paper instruments or a twenty-point scale when using a computerized version.

Participants saw the stimulus word and responded on each scale. The valence scale featured a smiling figure at one end (representing pleasantness) and a frowning figure at the other end (for unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at one end with a sleepy figure at the other, representing stimulating and unstimulating respectively. Finally, the control scale featured a large figure, indicating the highest degree of control, at one end and a small figure, indicating a lack of control, at the other end (Bradley \& Lang, 1999). The end result of this procedure yielded affective norms along the three dimensions for 1,040 English words (Bradley \& Lang, 1999). ANEW represented an important first step in establishing affective norms for large numbers of English words. However, later researchers found the 1,040-word list to be limiting for a language consisting of thousands of words.

Warriner, Kuperman, and Brysbaert (2013) exponentially lengthened the list of words with affective norms to 13,915 English lemmas, the base forms of words without inflection (i.e. \enquote{watch} rather than \enquote{watched} and \enquote{watching}). The researchers recognized the importance of affective norms in several areas of study, including emotion, language processing, and memory (Warriner et al., 2013). They argue the list of words included in ANEW is sufficient for small-scale factorial research designs, but the list is \enquote{prohibitively small} for larger-scale \enquote{megastudies} that are common in psycholinguistic research today (Warriner et al., 2013).

In order to source a large number of lemmas for affective ratings, the researchers drew from several validated sources. These include the 30,000 lemmas with age-of-acquisition (average age at which a particular word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US corpus consisting of subtitles from various forms of visual media (Warriner et al., 2013). This data collection resulted in the final list of 13,915 lemmas. Lists of 346-350 words were presented to participants recruited through the Amazon Mechanical Turk subject pool. Participants rated the words along one of the three dimensions, unlike the ANEW project in which participants rated each word along all three dimensions at once. The researchers used a nine-point scale similar to the one used by Bradley and Lang (1999) when collecting ratings for ANEW (Warriner et al., 2013).

The researchers noted several points of interest upon observing ratings. First, they found that valence and dominance ratings had a negative skew, indicating more words elicited feelings of happiness and control than their respective opposites. Also, when examining the relationship between valence and arousal ratings, the researchers found a U-shaped relationship. This indicates words with high degrees of positivity and negativity elicited higher arousal (Warriner et al., 2013). These observations along with the now-greatly expanded list of affective norms has been applied to several lines of inquiry in psycholinguistics.

Warriner and Kuperman (2015) utilized the new affective norms list in order to investigate the validity of the Pollyanna hypothesis, or the prevalence of a generally optimistic outlook in humans as reflected in language. The researchers were able to conclude the existence of a greater number of positive-valence English words in the list of 13,915 lemmas. Additionally, after observing token frequency in a number of text corpora, including SUBTLEX-US, COCA, BNC, TASA, and HAL, the researchers found that words with postie valence were also used more frequently (Warriner \& Kuperman, 2015). While the researchers concede the possibility of an acquiescence bias in ratings as a possible explanation for the observed positivity bias, this investigation represents one application of the Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner norms (2013) have been utilized in cognitive research as well. One cognition-based study investigates the relationship between emotion and response latencies in word recognition. Kuperman, Estes, Brysbaert, and Warriner (2014) sought to use the Warriner et al. (2013) norms to fill in the knowledge gaps regarding variance in word recognition. The researchers drew several conclusions regarding emotion and word recognition (specifically in naming and lexical decision tasks). First, Kuperman et al. (2014) found slower decision-making and reading times in negative-valence words, faster times in neutral words, and even faster times in words with positive valence. The researchers also concluded that words causing higher arousal tend to have slower decision times than less-arousing words. They found valence had a stronger effect on recognition than arousal (both effects were independent, not interactive). They found an interaction between emotion and word frequency such that valence and arousal are more effective on lower frequency words than high frequency words. Finally, Kuperman et al. (2014) found a greater effect of valence and arousal on response latency for lexical decision tasks than for naming tasks (Kuperman et al., 2014). This research serves as further evidence that the Warriner and Kuperman (2015) list can be used for research inquiries both within and without the field of psycholinguistics.

In the present studies, the researchers used the Warriner et al. (2013) list in order to denote the valence of the words appearing in the news articles scraped from the internet. Valence was considered as another independent variable and its relationship with the words comprising the Moral Foundations Dictionary were of chief interest to the researchers. The valence was used as a means to determine whether individual words in the MFD represented more positive aspects of their respective foundation or if they denoted a more negative aspect of the foundation. Incorporating word valence into a study involving the MFD is meant to alleviate some of the issues regarding the aforementioned ambiguity regarding the words in the Moral Foundations Dictionary.

@@

As in Experiment 1, the text was scanned using the \emph{ngram} package in \emph{R} at the conclusion of data collection (Schmidt et al., 2017). Using the package's word count function, articles that came through as blank text were removed, as well as text scraped from the Disqus commenting system. Duplicate articles were subsequently discarded.
